---
title: "Stat 133 HW03: Flow Control Structures and Functions with R"
author: "Hye Soo Choi and 23274190"
output: pdf_document
---

## Introduction

This assignment has two purposes:

a) to familiarize you with control flow structures in R 
b) to introduce you to writing functions in R

Submit your assignment to bcourses, specifically turn in your __Rmd__ (R markdown) file as well as the produced pdf file. Make sure to change the argument ```eval=TRUE``` inside every testing code chunk.

-----

## Last Element

Write a function ```last()``` that takes a vector (or factor) and returns the last element in the vector.
```{r}
# write your function
# last()
last <- function(vec){
  return (vec[length(vec)])
}

```

Test it:
```{r, eval = TRUE}
last(c('A', 'E', 'I', 'O', 'U'))
last(c(2, 4, 6, 8, 10))
```


-----

## If-then-else

Write a function ```multfive()``` that takes a number and determines whether the number is multiple of 5. If the provided number is multiple of five, then the output must be: ```it is multiple of five```. Conversely, if the provided number is not a multiple of five, then the output must be: ```it's not a multiple of 5```.
```{r}
# write your function
# multfive()
multfive <- function(num){
  if(num %% 5 == 0){
  return ('it is multiple of five')
  }
  else{
   return("it's not a multiple of 5")
  }
}

```

Test it:
```{r, eval=TRUE}
# multiple of five
multfive(10)

# not a multiple of five
multfive(33)
```


-----

## Create your histogram plotting function

Write a function ```histogram()``` that plots a histogram with added vertical lines for the following summary statistics: minimum value, median, mean, and maximum value. The main idea is to wrap the high-level function ```hist()``` and then plot the lines with a low-level plotting function.

Define your function with the following requirements: 

- bars of histogram colored in "gray90""
- line of minimum value in color "gray30", and dashed type
- line of maximum value in color "gray30", and dashed type
- line of median value in color "orange"
- line of mean value in color "tomato"
- all lines (min, max, median, mean) with a width of 3
```{r}
# write your function
# histogram
histogram <- function(data){
  hist(data, col = 'gray90', xlab = 'x', ylab ='Frequency',main='Histogram of x')
  abline(v =c(min(data), max(data), median(data), mean(data)), 
         col= c('gray30','gray30', 'orange','tomato'),lty = c(2,2,1,1), lwd=3)

}
```

Test it:
```{r, fig.width=5, fig.height=4, eval = TRUE}
histogram(mtcars$mpg)
```


-----

## Converting Fahrenheit Degrees

The table below shows the different formulas for converting Fahrenheit degrees into other scales:

| Units   | from Fahrenheit     |
| ------- | ------------------- |
| Celsius | (째F - 32) x 5/9     |
| Kelvin  | (째F + 459.67) x 5/9 |
| Reaumur | (째F - 32) x 4/9     |
| Rankine | 째F + 459.67         |

Write a function that converts from Fahrenheit degrees into each type of the four alternative scales. This implies writing four different functions:

- ```to_celsius()```
- ```to_kelvin()```
- ```to_reaumur()```
- ```to_rankine()```

```{r}
# write your functions
to_celsius <- function(faren){
  return ((faren - 32) * 5 / 9)
}
to_kelvin <- function(faren){
  return((faren + 459.67) * 5 / 9)
}
to_reaumur <- function(faren){
  return((faren - 32) * 4 / 9)
}
to_rankine <- function(faren){
  return(faren + 459.67)
}

```

Test them:
```{r, eval=TRUE}
to_celsius(34)
to_kelvin(34)
to_reaumur(34)
to_rankine(34)
```

-----

## Using switch()

Create a function ```convert()``` that converts Fahrenheit degrees into the specified scale. Use ```switch()``` and the previously defined functions---```to_celsius()```, ```to_kelvin()```, ```to_reaumur()``` and ```to_rankine()```---to define ```convert()```. Use two arguments: ```x``` and ```to```
```{r}
# write your function
# convert()
convert <- function( x, to ){
  scaled_temp <- switch(
    to,
    celsius = to_celsius(x),
    kelvin = to_kelvin(x),
    reaumur = to_reaumur(x),
    rankine = to_rankine(x)
  )
  return(scaled_temp)
}
```

Test it:
```{r, eval=TRUE}
convert(32, "celsius")
convert(32, "kelvin")
convert(32, "reaumur")
convert(32, "rankine")
```


-----

## Permutations

Write a function ```permute()``` that calculates the number of permutations of $k$ objects from a set of $n$ objects. 
```{r}

# write your function
# permute()

permute <- function(n,k){
  if (n < 0 | k < 0){
    print('stop: both inputs should be non-negative numbers')
    # I wanted this to be stop('both inputs should be non-negative numbers')
    # but when evaluating permute(-6, 6), this function stops and prevent pdf from being produced.
  }
  else if(n < k){
    return ( 0 )
  }
  else{
    n <- as.integer(n)
    k <- as.integer(k)
    return( factorial(n)/ factorial(n-k))
  }
}

```

Test it:
```{r, eval=TRUE}
permute(6, 2)
```

Make sure that the function checks that both $n$ are $k$ are non-negative numbers (if any of them is negative, the function must stop). Also make sure that if $n$ is less than $k$, the result is zero. In addition, $n$ and $k$ should be coerced as integers.
```{r, eval=TRUE}
# the following calls should not work
permute(2, 6)
permute(-6, 6)
```


-----

## Average function with for loop

Create a function ```average()``` using a ___for loop___ to compute the mean. ```average()``` takes a numeric vector and returns the average.
$$
\bar{x} = \frac{1}{n} \sum^{n}_{i = 1} x_i
$$
```{r}
# write your function
# average()
average <- function(num_vec){
  total <- 0
  for(i in 1:length(num_vec)){
    total <- total + num_vec[i]
  }
  return ( total / length(num_vec))
}

```

Test it:
```{r, eval=TRUE}
average(1:5)
mean(1:5)
```


-----

## Geometric Mean function

The formula of the geometric mean is:

$$
\left( \prod^{n}_{i = 1} x_i \right)^{1/n}
$$

Write a function ```geomean()``` that computes the geometric mean of a vector of positive numbers, using a ___for loop___:
```{r}
# write your function
# geomean()
  geomean <- function(num_vec){
    product <- 1
    for(i in 1:length(num_vec)){
      product <- product * num_vec[i]
    }
    return (product ^ (1/length(num_vec)))
  }
```

Test it:
```{r, eval=TRUE}
geomean(1)
geomean(1:5)
```


-----

## Frequency Table

Write a function ```freq_table()``` that takes a factor and generates a frequency table with 5 columns: 

1) ```category```: the levels of the factor
2) ```count```: absolute frequency
3) ```prop```: relative frequency (use four decimal places)
4) ```cumcount```: cumulative absolute frequency
5) ```cumprop```: cumulative relative frequency (use four decimal places)

Make sure that the input is a factor (otherwise the function should stop). Likewise, the output should be in ```data.frame``` form.

```{r}
# write your function
# freq_table()

freq_table <- function( data){
  if(is.factor(data) == FALSE){
    stop ('input should be a factor')
  }
  else{
  category <- levels(data)
  count <- as.vector(table(data))
  prop <- round(count/ length(data), digits = 4)
  cumcount <- cumsum(count)
  cumprop <- round (cumcount / length(data), digits = 4)
  }
  freq_table <- data.frame(category = category, count = count, prop = prop,
                           cumcount = cumcount, cumprop = cumprop)
  return (freq_table)
}
```

Test it:
```{r, eval=TRUE}
# some factor
set.seed(13)
sizes <- factor(
  sample(c('small', 'medium', 'large'), size = 90, replace = TRUE)
)

# frequency table
freq_table(sizes)
```


-----

## Summary Statistics Table

Write a function ```stats()``` that takes a numeric vector and generates the following descriptive statistics: 

- ```min```: minimum value
- ```max```: maximum value
- ```range```: range (max - min)
- ```q1```: first quartile
- ```q3```: third quartile
- ```iqr```: inter-quartile range (q3 - q1)
- ```median```: median
- ```mean```: mean
- ```sd```: standard deviation
- ```NAs```: number of missing values ```NA```

The function ```stats()``` should include an argument ```na.rm``` ---that takes a logical value--- so it can handle potential missing values. The output must be a data.frame of one column.

```{r}
# write your function
# stats()
stats <- function (num_vec , na.rm = FALSE){
  min <- min(num_vec, na.rm = na.rm)
  max <- max(num_vec, na.rm = na.rm)
  range <- max - min
  q1 <- quantile(num_vec, na.rm = na.rm)[2]
  q3 <- quantile(num_vec, na.rm = na.rm)[4]
  iqr <- q3 - q1
  median <- median(num_vec, na.rm = na.rm)
  mean <- mean(num_vec, na.rm = na.rm)
  sd <- sd(num_vec, na.rm = na.rm)
  NAs <- sum(is.na(num_vec))
  
  stats <- c( min, max,  range,  q1, q3,  iqr,  median,
             mean,  sd,  NAs)
  stats <- data.frame(stats)
  row.names(stats) <- c('min', 'max' , 'range', 'q1', 'q3' ,'iqr', 'median','mean', 'sd', 'NAs' )
  return( stats )
}
```

Test it:
```{r, eval=TRUE}
# no missing values
stats(1:10)
# missing values
stats(c(1:4, NA, 6:9, NA), na.rm = TRUE)
```


-----

## Frequency Table and Summary Statistics

Having created the functions ```freq_table()``` and ```stats()```, use them to write a function ```univarite()``` for producing summary statistics depending on the type of input. If the provided input is a numeric vector, then ```stats()``` should be called. In turn, if the provided input is a factor, then ```freq_table()``` should be called. If the input is not a numeric vector or a factor, then ```univariate()``` will print: ```"x must be either a numeric vector or a factor"```

```{r}
# write your function
# univariate()
univariate <- function(data){
  if(is.numeric(data)){
    return (stats(data))
  }
  else if(is.factor(data)){
    return (freq_table(data))
  }
  else{
    print ('x must be either a numeric vector or a factor')
  }
}
```

Test it:
```{r, eval=TRUE}
# factor input
univariate(sizes)

# numeric input
univariate(1:10)
```

This should not work:
```{r, eval=TRUE}
# this should cause an error
univariate(colors()[1:5])
```

