---
title: "Lab 15 Solutions"
author: "Donovan Lieu"
date: "August 10, 2015"
output: html_document
---

## SQL

This week, we'll be working with SQL, specifically with MySQL.
SQL (Structured Query Language) is a programming language intended to manipulate data in a relational database management system (RDBMS).
That is, it works with databases, each consisting of tables that can be worked with independently, or together.
Today, we'll mostly just be setting up everything and exploring very basic capabilities.
Tuesday and Wednesday, we'll do some more interesting things.

SQL is a declarative programming language, unlike R.
In SQL, instead of giving explicit instructions, the user provides a set of criteria for the result, but does not explicitly give instructions on how to obtain this result.
You may observe that this type of programming is exhibited by regex, in which a pattern is supplied to be matched, but no code is written to actually find the match.

### Beginning with MySQL

First of all, you'll want to download MySQL from the following website: "http://dev.mysql.com/downloads/mysql/".
You only need the MySQL Server, but feel free to install the entire package.

Make a note of your user name, if you specify it.
By default, there should be an account with user name "root" (I think).
Also, remember your password (you'll need it).

Once you've got the MySQL Server installed, locate the mysql application, and add it to your path.

Now we can run MySQL from the command line.
To start MySQL, run `mysql -u root -p`.
(Replace root with your user name if it is different.)
Then type your password.

The command `SHOW DATABASES;` shows all of the databases on your current MySQL directory.

You can create a database with the command `CREATE DATABASE db;`.
(Change "db" to the name of the database you want to create.)

To move inside a particular database, say `db`, use the command `USE db;`.

Once inside a database, the command `SHOW TABLES;` will list the tables in that database.

Later, we'll see how to actually work with databases and the tables inside them.

### MySQL in R

Several packages exist in R that allow a user to manipulate data in both R and SQL.
Since we're using MySQL, we'll be using the R package RMySQL.
Install this package.

```{r, eval = FALSE}
install.packages("RMySQL")
```

And load the package.

```{r, eval = FALSE}
library(RMySQL)
```

Once the package is loaded, we need to access the MySQL server.

```{r, eval = FALSE}
drv <- dbDriver("MySQL")
con <- dbConnect(drv, user = "root", dbname = "mysql", password = "s133",
  host = "localhost")
# Replace user, dbname, password, and host as necessary.
```

When you're done with MySQL, disconnect from the server and unload the driver as follows:

```{r, eval = FALSE}
# Don't run these yet!
dbDisconnect(con)
dbUnloadDriver(drv)
```

There are some specialized functions built into the RMySQL package, but they are mostly unnecessary.
For example, to list the tables in your current database:

```{r, eval = FALSE}
# Built-in RMySQL function
dbListTables(con)

# Or
query <- "SHOW TABLES;"
dbGetQuery(con, query)
```

See the documentation for other examples of such built-in functions.

Importantly, we can add data frames in R to a database in MySQL.

```{r, eval = FALSE}
dbDisconnect(con)
con <- dbConnect(drv, user = "root", dbname = "db", password = "s133",
  host = "localhost")

dbWriteTable(con, "state", as.data.frame(state.x77))
```

Now you can check that we have succesfully added a table to our database.

```{r, eval = FALSE}
dbGetQuery(con, "SHOW TABLES;")

# View contents of table
dbGetQuery(con, "SELECT * FROM state;")
# More on how to do this shortly
```

### Basic Subsetting in SQL

The most common command in SQL is the `SELECT` command, used to subset a table or tables.
For now, we'll work with only one table at a time.

The basic syntax is `SELECT select_expr FROM table`, where select_expr is a set of columns and table is the table we are subsetting.

For example:

```{r, eval = FALSE}
dbGetQuery(con, "SELECT row_names, Population FROM state;")
```

As you saw earlier, using `*` as the select expression will return all columns.

Finally, for now, notice that select expressions allow us to subset certain columns.
We can modify a select expression to only give us certain rows as well, using the `WHERE` command.

For example:

```{r, eval = FALSE}
dbGetQuery(con, "SELECT row_names FROM state WHERE Population >= 10000;")

# Equivalently, using `NOT` for negation
dbGetQuery(con, "SELECT row_names FROM state WHERE NOT Population < 10000;")
```

Multiple constraints can be imposed by using `AND` and `OR`.
Strings should be denoted using single quotes.

Your turn!

Extract the names of those states with an illiteracy rate of less than 1% and a murder rate of less than 10.
Hopefully, this should be fairly simple.

```{r, eval = FALSE}
# Fill in.
dbGetQuery(con, "SELECT row_names FROM state WHERE Illiteracy < 1 AND Murder < 10")
```

Don't forget to disconnect from the MySQL server and unload the driver!
