---
title: "Lab 16"
author: "Your Name"
date: "August 11, 2015"
output: html_document
---

## SQL (Continued)

We'll continue where we left off yesterday, and then we'll move on to some more SQL commands.

Before running MySQL, you'll need to first start the MySQL Server.
On Windows, run the Command Prompt as administrator, and use the command `mysqld`.
On Mac, go to System Preferences, find MySQL, and click "Start MySQL Server".

When you want to stop the server, you can use the command `mysqladmin -u root shutdown -p`, after which you should enter your server password.
If your server does not have a password, then omit the `-p`.

Continue with lab 15 before returning here ...

```{r, eval = FALSE}
library(RMySQL)

drv <- dbDriver("MySQL")
con <- dbConnect(drv, user = "root", dbname = "mysql", password = "",
  host = "localhost")

dbGetQuery(con, "USE db;")
```

### Aggregate Functions

SQL contains a few aggregate functions which operate on the rows of a given table.
These include: `COUNT`, `SUM`, `AVG`, `MIN`, and `MAX`.
Hopefully the purpose of these functions should be relatively obvious without explanation.

A simple example:

```{r, eval = FALSE}
dbGetQuery(con, "SELECT AVG(Population) FROM state;")
```

Recall from the previous lab that `*` returns all rows of the table.

Then predict what the following line does in general, and run it to check your prediction.

```{r, eval = FALSE}
# What should this evaluate to?
dbGetQuery(con, "SELECT COUNT(*) FROM state;")
```

The output of an SQL query can also be used as the input to another SQL query.

Your turn!

Determine which states have above average income, but below average high school graduation rates.

```{r, eval = FALSE}
# Fill in.
dbGetQuery(con, "SELECT row_names FROM state WHERE Income > (SELECT AVG(Income) from state) AND Illiteracy < (SELECT AVG(Illiteracy) from state);")
```

### Other Basic SQL Clauses

The `GROUP BY` clause can be appended to an SQL query.
This clause allows aggregate functions to be independently applied to subsets of a given table.

The `HAVING` clause can be further appended after a `GROUP BY` clause.
This clause allows the user to retain only those groups which satisfy a certain criterion (or criteria).

Your turn!

Consider the built-in dataset `sleep`.

```{r, eval = FALSE}
sleep
```

First add a table containing this dataset o the database `db`.
Then determine the mean, minimum, and maximum values of extra for the entire dataset.
Finally determine the mean, minimum, and maximum values of extra for each group.

```{r, eval = FALSE}
# Fill in.
```

When you're done with the lab, make sure to disconnect from the MySQL Server and unload the driver.

```{r, eval = FALSE}
dbDisconnect(con)
dbUnloadDriver(drv)
```
